---
title: "BAnalyzer - Barcode sequence analysis report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
theme: default
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}

knitr::opts_chunk$set(out.width = '80%',fig.asp= 0.5,fig.align='center',echo=FALSE, warning=FALSE, message=FALSE)
options(markdown.HTML.header = system.file("misc", "datatables.html", package = "knitr"))

library(DT, quietly = T)
library(tidyverse, quietly = T)
library(plotly, quietly = T)

executor <- Sys.info()["user"]
```

---
author: `r paste0(executor)`
---

```{r init}
workdir <- snakemake@params[["workdir"]]

distanceTable <- snakemake@input[["dist"]]
sizeHist <- snakemake@input[["sizedist"]]
nsequences <- snakemake@input[["seq"]]
ntaxids <- snakemake@input[["taxids"]]

trimming <- snakemake@params[["trimming"]]
database <- snakemake@params[["database"]]
identity <- snakemake@params[["id"]]
primers <- snakemake@params[["primers"]]

```

# Analysis report for database: `r basename(database)` {.tabset}

## Overview

* Database path: `r database`

* Number of sequences in database: `r scan(nsequences)`

* Number of taxids in database: `r scan(ntaxids)`

* Run directory: `r workdir`

* Primers were trimmed before alignment: `r trimming`

```{r primer_file, results='asis', eval=trimming}
cat("* Primer file: ", primers) 
```

## Pairwise distances


Pairwise comparison of sequences with less than `r identity*100`% identity was not performed.


```{r distances}
data_table <- read.csv(file = distanceTable, sep = '\t', check.names = FALSE)
datatable(data_table, filter = list(position = 'top'), rownames = FALSE, escape = FALSE,
		extensions= list("ColReorder" = NULL, "Buttons" = NULL),
		options = list(	
					dom = 'BRrlftpi',
					autoWidth=FALSE,
					scrollX = TRUE,
					lengthMenu = list(c(10, 50, 100, -1), c('10', '50', '100', 'All')),
					ColReorder = TRUE,
					buttons =
					list(
						'copy',
						'print',
						list(
						extend = 'collection',
						buttons = c('csv', 'excel', 'pdf'),
						text = 'Download'
						),
						I('colvis')
						)
						))
```

*Distance is given as the Hamming distance of the global pairwise alignment: each mismatch and gap adds one to the pairwise distance*

## Distance distribution

```{r dist_dist}
p <- ggplot(data_table, aes(x=distance)) +
		geom_histogram(binwidth = 1, color = 'black', fill = 'grey') +
		labs(title="Distance distribution", x="Hamming distance", y= "Count") +
		theme_classic()
		
ggplotly(p)
```

*This graph excludes self-comparisons and considers **both** orientations for each pair*

## Sequence lengths

```{r sizes}
size_table <- read.csv(sizeHist, sep = '\t', check.names = FALSE)
datatable(size_table, filter = list(position = 'top'), rownames = FALSE, escape = FALSE,
		extensions= list("ColReorder" = NULL, "Buttons" = NULL),
		options = list(	
					dom = 'BRrlftpi',
					autoWidth=FALSE,
					scrollX = TRUE,
					lengthMenu = list(c(10, 50, 100, -1), c('10', '50', '100', 'All')),
					ColReorder = TRUE,
					buttons =
					list(
						'copy',
						'print',
						list(
						extend = 'collection',
						buttons = c('csv', 'excel', 'pdf'),
						text = 'Download'
						),
						I('colvis')
						)
						))
```

## Sequence size distribution

```{r size_dist}

l <- htmltools::tagList()

if (trimming) {
	praw <- ggplot(size_table, aes(x=db_length)) +
			geom_histogram(binwidth = 1, color = 'black', fill = 'grey') +
			labs(title="Length distribution of original sequences", x="Length [bp]", y= "Count") +
			theme_classic()
	l[['praw']] <- ggplotly(praw)
		
	ptrim <- ggplot(size_table, aes(x=trim_length)) +
			geom_histogram(binwidth = 1, color = 'black', fill = 'grey') +
			labs(title="Length distribution of trimmed sequences", x="Length [bp]", y= "Count") +
			theme_classic()
	l[['ptrim']] <- ggplotly(ptrim)
} else {
	p <- ggplot(size_table, aes(x=length)) +
			geom_histogram(binwidth = 1, color = 'black', fill = 'grey') +
			labs(title="Length distribution of sequences", x="Length [bp]", y= "Count") +
			theme_classic()
	l[['p']] <- ggplotly(p)
}
l
```